---
title: "P Values - Exercises"
author: "Faiyaz Hasan"
date: "July 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1
---------
**Consider again the `mtcars` dataset. Use a two group t-test to test the hypothesis that the 4 and 6 cyl cars have the same `mpg`. Use a two sided test with unequal variances. Give a P-value.**

```{r solution 1}
library(datasets) # load datasets library
data("mtcars") # load mtcars dataset
str(mtcars) # show structure of dataset
fourcyldata <- mtcars[mtcars$cyl == 4, ] # subset of 4 cylinder cars
sixcyldata <- mtcars[mtcars$cyl == 6, ] # subset of 6 cylinder cars
hypothtest <- t.test(fourcyldata$mpg, sixcyldata$mpg, alternative = "two.sided", var.equal = FALSE) # two sided t test with unequal variances
```

Hence, the two sided t-test with unequal variances take a p-value of `r hypothtest$p.value`.

Problem 2
---------
**A website was monitored for a year and it recieved 520 hits per day. In the first 30 days in the next year, the site recieved 15800 hits. Assuming that web hits are Poisson, give an exact one sided P-value to the alternate hypothesis that webhits are up this year over last. Do you reject the hypothesis?**

```{r solution 2}
pval <- ppois(15799, lambda = 520 * 30, lower.tail = FALSE) # P-value
```

Hence, the p value is given by `r pval` and we do not reject the null hypothesis.

Problem 3
---------
**Consider the `mtcars` data set: (i) Give the p-value for a t-test comparing MPG for 6 and 8 cylinder cars assuming equal variances, (ii) Give the associated p-value for a z-test, (iii) Give the common standard deviation estimate for MPG across cylinders, (iv) Would the t-test reject at the two sided 0.05 level?**

```{r solution 3 (i)}
library(datasets) # load datasets library
data("mtcars") # load mtcars dataset
str(mtcars) # show structure of dataset
sixcyldata <- mtcars[mtcars$cyl == 6, ] # subset of 6 cylinder cars
eightcyldata <- mtcars[mtcars$cyl == 8, ] # subset of 8 cylinder cars
ptval <- t.test(sixcyldata$mpg, eightcyldata$mpg, paired = FALSE, var.equal = TRUE, alternative = "two.sided")$p.value # t-test p value
```
The p-value from the t-test is `r ptval`.


```{r sol 3 (ii)}
# compute z statistic
m1 <- mean(sixcyldata$mpg) # mean of 6 cyl. data
m2 <- mean(eightcyldata$mpg) # mean of 8 cyl. data
sd1 <- sd(sixcyldata$mpg) # standard deviation of 6 cyl. data 
sd2 <- sd(eightcyldata$mpg) # std dev of 8 cyl data
n1 <- length(sixcyldata$mpg) # number of 6 cyl observations
n2 <- length(eightcyldata$mpg) # number of 8 cyl observations
sdcom <- sqrt(((n1 - 1)*sd1^2 + (n2-2)*sd2^2)/(n1+n2-2))
se <- sdcom*sqrt(1/n1 + 1/n2) # standrd error
zstat <- (m1-m2)/se # corresponding z statistics
pzval <- 2*pnorm(-abs(zstat)) # z-test p value
```
The p-value from the z-test is `r pzval`.

The common standard error is `r sdcom`. We would reject the null hypothesis that the MPG of 6 and 8 cylinder cars are equivalent.