---
title: "Hypothesis Testing - Exercises"
author: "Faiyaz Hasan"
date: "July 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1
---------
**Load the data set `mtcars` in the `datasets` R package. Assume that the data set mtcars is a random sample. Compute the mean MPG, $\bar{x}$, of this sample. You want to test whether the true MPG is $\mu_0$ or smaller using a one sided 5% level test. ($H_0 : \mu = \mu_0$ versus $H_a : \mu < \mu_0$). Using the data set and a Z test based on the mean MPG of the sample, what is the smallest value of $\mu_0$ that you would reject for (to two decimal places)?**

```{r solution 1}
library(datasets)
data(mtcars) # load mtcars data set
str(mtcars) # structure of data set
meanmpg <- mean(mtcars$mpg) # [mean] of [m]iles [p]er [g]allon of data set
sdmpg <- sd(mtcars$mpg) # [s]tandard [d]eviation of [mpg]
n <- length(mtcars$mpg) # length of data set, i.e. sample size
meanmpg
smallestvalue <- meanmpg + qnorm(0.95)*sdmpg/sqrt(n) # Solution
```
Hence, the smallest value of $\mu_0$ based on the Z test which would lead us to reject $H_0$ is `r smallestvalue`.

Problem 2
---------
**Consider again the mtcars dataset. Use a two group t-test to test the hypothesis that the 4 and 6 cylinder cars have the same mpg. Use a two sided test with unequal variances. Do you reject?**

```{r solution 2}
fourcyldata <- mtcars[mtcars$cyl == 4, ] # dataset of 4 cylinder cars only
sixcyldata <- mtcars[mtcars$cyl == 6, ] # dataset of 6 cylinder cars

# Carry out the T test that 4 and 6 cyl. cars have same mpg.
t.test(fourcyldata$mpg, sixcyldata$mpg, paired = FALSE, var.equal = FALSE)
```

At the 5% level, we reject the hypothesis.

Problem 3
---------
**A sample of 100 men yielded an average PSA level of 3.0 with a standard deviation of 1.1. What are the complete set of values that a 5% two sided Z test of $H_0: \mu = \mu_0$ would fail to reject the null hypothesis for?**

```{r solution 3}
# range of values
valuerange <- 3 + c(-1, 1)*qnorm(0.975)*1.1/sqrt(100)
```

We would fail to reject the hypothesis in the interval `r valuerange`.


