---
title: "T confidence intervals - Exercises"
author: "Faiyaz Hasan"
date: "July 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Problem 1  
---------
**Load the data set mtcars in the `datasets` R package. Calculate a 95% confidence interval to the nearest mile per gallon (MPG) for the variable `mpg`.**  

```{r load mtcars dataset}
setwd("C:\\Users\\Windows\\Documents\\JHU_Data_Science\\Course_6\\") # Set working directory
library(datasets) # Load datasets package  
data(mtcars) # Load mtcars data  
head(mtcars) # Heading of mtcars dataset
```
Now, let us find the 95% confidence interval for the mpg variable.

```{r mpg mean and sd}
meanmpg <- mean(mtcars$mpg) # mean of mpg data
sdmpg <- sd(mtcars$mpg) # standard deviation of data
n <- length(mtcars$mpg) # sample size

meanmpg + c(-1, 1)*qt(0.975, n-1)*sdmpg/sqrt(n) # Solution 1 - direct computaion

t.test(mtcars$mpg) # Solution 2 - Use predefined R function
```
Hence, the 95% confidence interval assuming a t distribution (sample size is 32) is `[18, 22]`.

Problem 2
---------  
**Suppose that the standard deviation of 9 paired differences is `1`. What value would the average of the difference have to be so that the lower endpoint of a 95% student's t confidence interval touches zero?**

The value of the average would have to be: 
```{r mean value computation}
qt(0.975, 8)*1/sqrt(9)
```

Problem 3
---------
**Consider the mtcars dataset. Construct a 95% T interval for the MPG comparing 4 to 6 cylinder cars (subtracting in the order of 4 - 6) assume a constant variance.**

```{r comparison of 4 and 6 cylinder cars}
head(mtcars)
fourcyldata <- mtcars[mtcars$cyl == 4, ] # dataset of 4 cylinder cars only
sixcyldata <- mtcars[mtcars$cyl == 6, ] # dataset of 6 cylinder cars
meanfour <- mean(fourcyldata$mpg) # mean of 4 cyl. data
meansix <- mean(sixcyldata$mpg) # mean of 6 cyl. data
sdfour <- sd(fourcyldata$mpg) # standard dev. of 4 cyl data
sdsix <- sd(sixcyldata$mpg) # standard dev. of 6 cyl data
nfour <- length(fourcyldata$mpg) # length of four cyl. data
nsix <- length(sixcyldata$mpg) # length of six cyl. data

# Solution 1 using formulae
(meanfour - meansix) + c(-1, 1)*sqrt(1/nfour + 1/nsix)*qt(0.975, nfour+nsix-2)*sqrt(((nfour-1)*sdfour^2 + (nsix-1)*sdsix^2)/(nfour+nsix-2))

# Solution 2 using R function
t.test(fourcyldata$mpg, sixcyldata$mpg, paired = FALSE, var.equal = TRUE)$conf
```


